---
title: "Senior Project Start"
author: "Michael Streyle"
date: "September 28th, 2018"
output: word_document
editor_options: 
  chunk_output_type: console
---
  
```{r echo=T}
```

# Beginning

```{r echo=F}

setwd("C:/Users/Michael Streyle/Desktop/Senior Project") #laptop
#setwd("C:/Users/Michael/Desktop/Senior Project") #desktop

data1  <- read.csv('Dataset_spine.csv', col.names = c('pelvic_incidence', 'pelvic_tilt', 'lumbar_lordosis_angle', 'sacral_slope', 'pelvic_radius', 'degree_spondylolisthesis', 'pelvic_slope', 'Direct_tilt', 'thoracic_slope', 'cervical_tilt', 'sacrum_angle', 'scoliosis_slope', 'classification', ' '))


data1$X. <- NULL #dropping the column with variable descriptions in it
par (mfrow = c(1,1))


data = scale(data1[, 1:12]) #scaling all except classification variable
data = data.frame(data)
data$classification = data1$classification #add classification back into scaled dataframe
data$class = ifelse(data$classification == "Abnormal", 0, 1) #making classification numeric

#these are the packages I use:

# library(ModelMetrics) old use for confusion matrices. wasnt working for factor columns
library(e1071) #SVM
library(randomForest)
library(caret) #for confusion matrices

```
# Introduction

The data used in this analysis is a collection of 310 observations of spinal measurements. The objective is to classify each observation as normal or abnormal using machine learning modeling. As seen in the pairs plot, the last 6 variables are essentially random noise, so they will be left out of the initial models. Later, I will use these random variabels to compare how the models handle extra noise. 

```{r echo=F}

attach(data)

```

# Initial Exploration

A histogram of each of the first 6 predictor variables reveals that they are each relatively normal. A histogram of the other 6 predictor variables confirms that they are not normally distributed but rather somewhat uniformly distributed. This aligns with the idea that the last 6 are random variables. There are no missing values in this data. 

```{r, fig.height=10, fig.width=10}
my_cols <- c("#00AFBB", "#E7B800", "#FC4E07")

pairs(data, col = my_cols[data$classification]) #pairs plot




```


```{r echo=F}
#histograms
hist(pelvic_incidence)
hist(pelvic_tilt)
hist(lumbar_lordosis_angle)
hist(sacral_slope)
hist(pelvic_radius)
hist(degree_spondylolisthesis)
hist(pelvic_slope)
hist(Direct_tilt)


#boxplots
boxplot(pelvic_incidence)
boxplot(pelvic_tilt)
boxplot(lumbar_lordosis_angle)
boxplot(sacral_slope)
boxplot(pelvic_radius)
boxplot(degree_spondylolisthesis)



```

